%!TEX root = ../sandReportSpec.tex


%\lstMakeShortInline[style=CppCodeInlineStyle]{\|}
\section{Data Access Handles}
\label{sec:handles}
\codetarget{AccessHandle<T>} objects are lightweight wrappers around the actual
data structure of interest having type \inlinecode{T}. 
The \glspl{handle} add a control block (metadata) that tracks uses of the \gls{handle} and enforces
\gls{sequential semantics}, analogous to smart pointers that wrap pointer types and provide a reference counting control block.
Critically, this interface is non-intrusive, wrapping any type \inlinecode{T} without requiring that type to be modified.

Most critically, \ahandle enables deferred access since \ahandle can exist in
\gls{ready} or \gls{pending} states. 
\Gls{ready} and \gls{pending} are not rigorously defined in the state table for \ahandle in
Section~\ref{sec:handlerules}, but rather guiding concepts.
\Gls{ready} \glspl{handle} can be dereferenced (have their underlying values
\gls{fetch}ed) and be used immediately to perform work.
\Gls{pending} \glspl{handle} cannot be dereferenced, but can still be used to \emph{schedule} or \emph{instantiate} work.
Thus even if a \gls{handle} is carrying \gls{pending} or unresolved data, execution can advance thereby unrolling more of the task graph.
This \emph{lookahead} is the key element that enables runtime optimizations.
Lookahead gives the \gls{runtime system} more complete knowledge of the task graph instead of locally executing step-by-step.
By looking ahead, the \gls{runtime system} can reorder or migrate \glspl{task} to maximize data locality and improve load balance.
The most critical conceptual change from standard \CC{} to \gls{DARMA} are
\gls{pending} variables that enable lookahead, unlike conventional \CC{}
variables that must always be ``\gls{ready}''.

\Glspl{handle} can be created three different ways.
\begin{enumerate}
  \item a \gls{handle} to data that does not yet exist in the system
but needs to be created, or
\item a \gls{handle} to data produced by another process that needs to
be read, or 
\item a \gls{handle} to data produced by another process that needs to be overwritten or modified. 
  Note that this type of \gls{handle} does not exist in the current version of
  the specification.
\end{enumerate}

Type 1 is denoted as \codelink{initial_access} in \gls{DARMA}, 
which informs the \gls{runtime system} that the data with the 
specified \codelink{key} does not yet exist, and the user intends to 
create this data.% and (potentially) publish it.
Hence, an \codelink{initial_access} data \gls{handle} is usually 
followed by a memory allocation, and a value assignment. 
Remark: Although we could explicitly write out \ahandleT in the code below,
we \emph{strongly encourage} programmers to use the \CC{} \inlinecode{auto} keyword.
It will greatly increase code portability for future (potentially
backwards-incompatible) versions of \gls{DARMA},
with the additional benefit of decreasing code verbosity.

\begin{CppCode}
auto float_handle = initial_access<float>("float_key");
create_work([=]{
  float_handle.set_value(3.14);
});
\end{CppCode}
As stated above, \gls{DARMA} provides two methods for expressing \gls{task}
\glspl{precondition}: \gls{sequential semantics} and \gls{coordination
semantics}.  \codelink{initial_access} is necessary in both methods.
Once created, a \gls{handle} can be passed along to \glspl{subtask} within the same,
sequential \gls{execution stream}.
A \gls{handle} created by \codelink{initial_access} can also be \codelink{publish}ed,
making it available to other \glspl{execution stream} via \gls{key-value store} coordination.

\Gls{handle} of Type 2 above request read-only access to data produced via
external \glspl{execution stream} through \codelink{read_access} (which causes
a \gls{fetch} to be performed). 
As such, \codelink{read_acccess} is only relevant for applications that use coordination to express data flow.
\begin{CppCode}
auto float_handle = read_access<float>("another_float_key");
create_work([=]{
  float val = float_handle.get_value();
  std::cout << "Value read with key another_float_key is " << val;  
});
\end{CppCode}

Immediately following the \codelink{read_access} function, the \ahandle will be
\emph{\gls{pending}} instead of \emph{\gls{ready}}.
To enable \codelink{get_value} to be called (put the \gls{handle} in a
\gls{ready} state), the \gls{handle} must be used inside of a \codelink{create_work}.
\cwork defers execution of the code block until the \gls{key-value store}
resolves the value of \inlinecode{float_handle} and converts it to a
\gls{ready} state.  This might involve moving data if the \inlinecode{float} is on a remote node.
Remark: future versions of \gls{DARMA} will enable \glspl{task} to begin
optimistically with some \glspl{handle} still in a \gls{pending} state, but
this is not supported in the current version of the specification.

In general, any calls to \codelink{get_value} should occur 
within a scoped code block to avoid dangling references to stale physical memory locations.
%Even when it becomes possible to wait on the data associated with a handle (in
%later versions of the spec), best practice is to access data via deferred tasks
inside a \codelink{create_work} block when possible.
\todo{Jeremy/David: this dangling sentence is part of a partially commented out
sentences - please fix.}




\subsection{Publish}
\label{subsec:publish}
By default, unless explicitly published, data \glspl{handle} are visible only
to \glspl{task} within the same scope (\glspl{task} that have a copy of the actual
\ahandleT object, created as discussed in Section~\ref{sec:handles}).
For data to be globally visible in the global memory space (\gls{key-value store}),
the application developer must explicitly \codelink{publish} data.  
Unpublished data will be reclaimed once the last \gls{handle} referencing it goes out of scope (i.e refcount goes to zero),
freeing the memory and resolving any anti-dependencies analogous to the
destructor invocation in \CC{} when a class goes out of scope.  
Unpublished data can leverage the \gls{sequential semantics} of the application
for \gls{garbage collection}. \todo[inline]{Jeremy/David: can you update the
  previous sentence.  Is it that the translation layer leverages sequential
  semantics for garbage collection? Or, does the runtime system handle it?
Either way the sentence doesn't quite make sense to me as is.}  
Published data, however, is globally visible to all \glspl{execution stream} and requires more ``permanence.''  
In order to reclaim the \todo{Jeremy/David, do you mean reclaim the memory
associated with the data here?} data (garbage collect or resolve
anti-dependence), published data must know its \gls{access group}.
When all read \glspl{handle} within an \gls{access group} have been deleted or
released \emph{globally}, the published \todo{Jeremy/David: again here, do you mean the
memory associated with the data?} data can be reclaimed.
The easiest way to declare an \gls{access group} (and currently the only supported method) is
to simply give the total number of additional read \ahandleT objects that will
be created referring to it (recall that read \ahandleT objects cause a
\gls{fetch} to be performed).
In future versions, hints will be supported about which specific \glspl{task} will need to read data.
This \codelink{publish}/\gls{fetch} mechanism replaces an analogous
\inlinecode{MPI_Send/Recv} or even \todo{Jeremy/David: why the use of
potentially here? If it supports it, then leave it in. If not, then can we add
something more descriptive as to why just potentially?} potentially an \inlinecode{MPI_Bcast}.
In MPI, these function calls force an \inlinecode{MPI_Send} or
\inlinecode{MPI_Wait} to block until the \gls{runtime system} guarantees that the data has been delivered.
An \gls{access group} in \gls{DARMA} provides a similar guarantee.
Until all readers in an \gls{access group} have received or released their
data, \gls{DARMA} cannot reclaim (garbage collect, clear anti-dependencies).

\begin{CppCode}
auto float_handle = initial_access<float>("float_key");
create_work([=]{
  float_handle.set_value(3.14);
});
float_handle.publish(n_readers=1);
\end{CppCode}
The \codelink{n_readers} specification in the \codelink{publish} call is a
\gls{keyword argument} (see Section~\ref{sec:keyword}) that informs the
\gls{runtime system} that the data
(associated with \inlinecode{float_key}) will only ever be read once, and hence can be safely garbage collected soon after.  
This code provides similar functionality to an MPI \inlinecode{send/receive}.

As discussed above, \glspl{handle} can either be \gls{ready} or \gls{pending}. 
In reality, the distinction is more subtle. 
The ``readiness'' and ``pendingness'' can be different for read usages and write usages.
Thus a handle can be read-ready, but modify-pending.
This will be the case after \codelink{publish} operations.
\Gls{publish} operations are treated as asynchronous read operations --- that is,
\inlinecode{h.publish(...)} is equivalent to
\begin{CppCode}
create_work(reads(h), [=]{...});  
\end{CppCode}
This means that the same precautions should be taken as with asynchronous reads. 
In particular, even if the handle was ready for modifying before
\codelink{publish}, it is no longer valid to call \inlinecode{h.set_value()}
after the \codlink{publish}.
The asynchronous read done by the \codelink{publish} may or may not have occurred yet.  
In this scenario, one should use instead
\begin{CppCode}
create_work([=]{ h.set_value(...); });
\end{CppCode}
to force the \gls{handle} from a \gls{pending} state to a \gls{ready} state.

\subsubsection{Publication Versions}
If a \gls{handle} is going to be published multiple times (or, more specifically, if
the \codelink{key} with which the \gls{handle} was created is going to be published multiple
times), it needs to be published with a different \codelink{version} each time.  
A \codelink{version} is just like a \codelink{key} --- an arbitrary \gls{tuple} of values (see Section~\ref{subsec:keys}).  
For instance:

\begin{minipage}{0.5\textwidth}
\begin{CppCode}
/** Execution stream 0 */
auto float_h =
  initial_access<float>("float_key");
auto int_h =
  initial_access<int>("int_key");
/* Execution stream 1 */
int_h.publish(n_readers=3, version=77);
//Use version() for multiple parts
float_ha.publish(n_readers=1, 
    version("alpha",42));
\end{CppCode}
\end{minipage}
\begin{minipage}{0.45\textwidth}
\begin{CppCode}
/* Execution stream 1*/
auto my_int =  read_access<int>(
  "int_key", version=77);
auto my_float = read_access<float>(
  "float_key", version("alpha",42));
\end{CppCode}
\end{minipage}

A \codelink{version} has similarities with an MPI tag, as they both ensure the uniqueness of data.
However, unlike MPI which uses a combination of message order and tag to
uniquely identify messages and match \inlinecode{send/recv} pairs,
the \gls{DARMA} asynchronous model does not allow implicit publication order to
be used in matching \gls{publish}/\gls{fetch} pairs.
Instead, all publications must uniquely identify each publication with a
specific \codelink{version}.


\subsection{Keys}
\label{subsec:keys}
In the examples in this section, the \codetarget{key} to the
\ahandleT has always been a single string. 
A \codelink{key} in DARMA 
can be an arbitrary \gls{tuple} of values.  This 
makes it very easy for the application developer to create an expressive
and descriptive \codelink{key} for each piece of data.  \Glspl{tuple} can comprise
different bit-wise copiable data types.  The example at the end of
Section~\ref{sec:spmd} illustrates the use of the \gls{rank} within the
\gls{handle} \codelink{key}.  The following example shows the use of an
aribitrary \gls{tuple} as a \codelink{key}:
\begin{CppCode}
  int neighbor_id
  double other_identifier;

  // some code that sets neighborID and other_identifier
  
  auto float_handle = initial_access<float>("float_key", 
                                            neighbor_id, 
                                            other_identifier);
\end{CppCode}

\subsection{Handle Usage Rules}
\label{sec:handlerules}
As alluded to above, \glspl{handle} are assigned states, and these states change 
based on the operations applied to them.  
The state of a \gls{handle} encompasses both its read/write permissions and its ``readiness."
\Gls{pending} \glspl{handle} can only be used for scheduling \glspl{task} while
\gls{ready} \glspl{handle} can be immediately used to do work.
Here we more rigorously divide permissions into two main categories: 
\begin{itemize}
\item[a] {\it Scheduling}: Permissions a handle may use when instantiating tasks with \inlinecode{create_work}.
These permissions apply independent of handle readiness. 
Generally, this will be Read (handle may only used in read-only tasks) or Modify (handle may be used in read-only or read-write tasks). 
\item[b] {\it Immediate}: Permissions that apply immediately, indicating the ``readiness'' of the handle.  Immediate privileges can ever be greater than Scheduling privileges.
A task can never have greater permissions doing immediate work than it can for instantiating deferred work.
\end{itemize}

For the two methods of creating handles, we have the following initializations.
\begin{itemize}
\item \inlinecode{initial_access<T>}: Initialized with scheduling modify, immediate none. 
The handle can be used in any mode when instantiating deferred work.
However, the handle is not necessarily initialized and as such cannot be used immediately for reads or writes.
%
\item \inlinecode{read_access<T>}: Initialized with scheduling read, immediate none.
The handle can only be used for reads when instantiating deferred work.
However, the handle is not necessarily initialized and as such cannot be used immediately.
\end{itemize}


In the parent task, a handle will have an initial pair of scheduling/immediate privileges (predecessor state).
After the call to create work, the handle's state will have changed, potentially losing some immediate privileges (continuing state).
As specified currently, execution does not begin inside the deferred work block until a handle becomes ready.
Inside the \cwork (capture state), the handle's immediate and scheduling privileges will therefore be the same.

Consider the following example:
\begin{CppCode}
auto float_handle = initial_access<float>("yet_another_float_key");
....
create_work(reads(float_handle), [=] {
  std::cout << "Value read with key yet_another_float_key is " 
          << float_handle.get_value() << std::endl;
})
create_work(reads(float_handle), [=] {
  float val = float_handle.get_value();
  if (val > 0) std::out << "Value is positive" << std::end;
})
//read-write work down here
\end{CppCode}
In this case, subtasks are created that only need read access. 
Without the \codelink{reads} qualifier, these \glspl{task} could not run in parallel (or out-of-order) since they
would by default request read-write privileges.
\Gls{sequential semantics} would then require them to write in-order sequentially.
This example highlights the importance that \glspl{task} only ever request the privileges they need. 
Over-requesting privileges will limit the amount of available parallelism in the code.
%A \gls{task} may be using a \gls{handle} with read-write privileges created by
%\codelink{initial_access}
%and then create a subtask that requires only read privileges.
Table \ref{tab:immsimp} summarizes the state transitions involving these three handles following \cwork.
\begin{CppCode}
//Predecessor parent state
create_work([=]{
  //Capture state
})
//Continuing parent state
\end{CppCode}

\begin{table}[!t]
\begin{center}
{\small
\begin{tabular}{cc|cc|cc}
 \hline
 \multicolumn{2}{c|}{Predecessor State}
 & \multicolumn{2}{c|}{\inlinecode{get_value()}} 
 & \multicolumn{2}{c}
 {
    \specialcell{ \inlinecode{emplace_value()} \\ 
                  \inlinecode{set_value()}\\
                  \inlinecode{get_reference()}\\
                  } 
 } 
  \\
 \hline
 \specialcell{Scheduling\\ permissions} 
 & \specialcell{Immediate\\ permissions}  
 & { {\footnotesize Allowed? } } \hspace{-0.cm} & { {\footnotesize Continuing as}}
 & { {\footnotesize Allowed? } } \hspace{-0.cm} & { {\footnotesize Continuing as}}
 \\
 \hline
 None & None
 & No & -
 & No & -
 \\
 %
 Read & None
 & No & -
 & No & -
 \\
 %
 Read & Read
 & Yes & {\em Read/Read}
 & No & -
  \\
 %
 Modify & None
 & No & -
 & No & -
 \\
 %
 Modify & Read
 & Yes & {\em Modify/Read}  
 & No & -
  \\
 %
 Modify & Modify
 & Yes & {\em Modify/Modify}  
 & Yes & {\em Modify/Modify}  
  \\
\hline
\end{tabular}
}
\caption{Operations on the various states}
\label{tab:immsimp}
\end{center}
\end{table}
%
\begin{table}[!t]
\begin{center}
{\small
\begin{tabular}{cc|ccc|ccc}
 \hline
 \multicolumn{2}{c|}{Predecessor State} 
 & \multicolumn{3}{c|}{\em{read-only capture} and publish} 
 & \multicolumn{3}{c}{\em{modify capture}}  \\
 \hline
 \specialcell{Scheduling\\ permissions} 
 & \specialcell{Immediate\\ permissions}  
 & {\footnotesize Allowed? } &  \specialcell{\footnotesize Capture\\Handle} 
 &  \specialcell{\footnotesize Continuing\\Handle} 
 & {\footnotesize Allowed? } &  \specialcell{\footnotesize Capture\\Handle} 
 &  \specialcell{\footnotesize Continuing\\Handle} \\
 \hline
 None & None & No & - & - & No & - & - \\
 %
 Read & None 
 & Yes 
 & {\em Read/Read}
 & {\em Read/None}
 & No
 & -
 & - \\
 %
 Read & Read
 & Yes 
 & {\em Read/Read}
 & {\em Read/Read}
 & No
 & -
 & - \\
 %
 Modify & None
 & Yes 
 & {\em Read/Read}
 & {\em Modify/None}
 & Yes
 & {\em Modify/Modify} 
 & {\em Modify/None} \\
 %
 Modify & Read
 & Yes 
 & {\em Read/Read}
 & {\em Modify/Read}
 & Yes
 & {\em Modify/Modify} 
 & {\em Modify/None} \\
 %
 Modify & Modify
 & Yes 
 & {\em Read/Read} 
 & {\em Modify/Read} 
 & Yes
 & {\em Modify/Modify}
 & {\em Modify/None} \\
\end{tabular}
}
\caption{Deferred (capturing) operations on the various states.}
\label{tab:capsimp}
\end{center}
\end{table}


To further illustrate, below is an \emph{incorrect} usage of modify privileges

\begin{minipage}[t]{0.45\linewidth}
\centering WRONG
\begin{vaspPseudo}
initial_access<int> a
//a is in Modify/None
a.set_value(1) $\no$ 
a.get_value()  $\no$
\end{vaspPseudo}
\end{minipage}
\begin{minipage}[t]{0.45\linewidth}
\centering CORRECT
\begin{vaspPseudo}
initial_access<int> a
//a is in Modify/None
create_work([=]{ //modify capture
    a.emplace_value(1)  $\yes$
    a.set_value(1)      $\yes$
    a.get_reference()=1 $\yes$
});
\end{vaspPseudo}
\end{minipage}


Additionally, we demonstrate an \emph{incorrect} usage of read privileges

\begin{minipage}[t]{0.45\linewidth}
\centering
\centering WRONG
\begin{vaspPseudo}
read_access<int> b
//b is in Read/None
b.get_value()   $\no$
b.set_value(1)  $\no$
create_work([=]{ //capture
  b.set_value(1) $\no$
});
\end{vaspPseudo}
\end{minipage}
\begin{minipage}[t]{0.45\linewidth}
\centering
\centering CORRECT
\begin{vaspPseudo}
read_access<int> b
//b is in Read/None
create_work([=]{ // capture
  b.get_value()  $\yes$
});
\end{vaspPseudo}
\end{minipage}



