%!TEX root = sandReportSpec.tex

<<<<<<< HEAD
In this document, we provide the specification for \gls{DARMA}, 
a co-design research vehicle for \gls{AMT} \glspl{programming model} and
\glspl{runtime system}.  \gls{DARMA} serves to:
1) insulate applications from \gls{runtime system} and hardware idiosyncrasies,
2) improve \gls{AMT} runtime programmability by co-designing an \gls{API} directly with
  application developers, 3) synthesize application \gls{co-design} activities into
  meaningful requirements for runtimes, and 4) 
facilitate \gls{AMT} design space characterization and definition, accelerating the
development of \gls{AMT} best practices.

The \gls{DARMA} code base comprises three main layers: the \gls{front end} \gls{API},
\gls{translation layer}, and \gls{back end} \gls{API}.  
The \gls{front end} is the user-level \gls{API}. 
Its use has the feel of an \gls{EDSL} in \CC,  inheriting the generic
language constructs of \CC and adding \gls{semantics} that facilitate
distributed parallel programming. Though the \gls{front end} \gls{EDSL} uses
\CC constructs in non-traditional ways to implement these semantics, it is
nonetheless fully embedded in the \CC language and
requires a widely supported subset of \CC{}14 functionality \compilerReqs.
The \gls{translation layer} leverages \CC\ \gls{template
metaprogramming} to map the user's code onto the \gls{back end} \gls{API}.
The \gls{back end} \gls{API} is a set of abstract classes and function
signatures that \gls{runtime system} developers must implement in accordance with the
specification requirements in order to interface with application code written
to the \gls{DARMA} front end.  
Each of these three layers is described herein. We also
include a list of application requirements driving the specification (along
with a list of the applications contributing to the requirements to date), some basic rules 
and examples, and conclude with a brief history of changes between previous
versions of the specification, and summary of the planned changes in upcoming
versions of the specification.  Appendices walk the user through a more detailed set
of examples of applications written in the \gls{DARMA} \gls{API} and provide
additional technical details for those who are interested.
=======
In this document, we provide the specification for DARMA 
\todo{acronym DARMA not defined anywhere, this is first time it appears},
a co-design research vehicle for \gls{AMT} programming models and runtime
systems.  This specification serves to facilitate three primary activities: 
1) gathering and communicating application requirements, 2) exploring AMT
design space tradeoffs, and 3) communication between the AMT community and
other computer science research communities.
As part of this document, we discuss application requirements driving the specification (along
with a list of the applications contributing to the requirements to date).

The DARMA code base comprises two main layers: 
an API for writing application code that expresses the algorithm and  
a translation layer implicitly invoked by application-level code that maps application-level constructs
to an abstract backend API.
Executable DARMA applications link to a third layer implementing the abstract backend runtime API.
The implementations may be external, but default implementations are provided in the DARMA code distribution.
The application-facing user-level \gls{API} has the feel of an \gls{EDSL} in \CC,  
inheriting the generic language constructs of \CC and adding \gls{semantics} that facilitate
distributed parallel programming.  
Though the \gls{EDSL} uses modern \CC constructs unfamiliar to many programmers to implement these semantics, 
it is nonetheless fully embedded in the \CC language and as such will compile with any
\CC compiler that supports the full \CC11 standard and subsets of \CC14.
There advanced C++ features are widely supported (\inlinecode{gcc >= 4.8, clang >= 3.5, icc >= 16}).  
The \gls{translation layer} leverages \CC\ \gls{template
metaprogramming} to map the user's code onto the \gls{back end} runtime \gls{API}.
Each of these three layers is described herein.


>>>>>>> darma/0.3-devel
