%!TEX root = sandReportSpec.tex

\chapter{Requirements} 
\label{chap:requirements} 
\todo[inline]{flesh out these sections}
Here we enumerate requirements gathered to date: 
\begin{compactitem}
\item Applications need an abstraction layer between applications and the
runtime system/physical architecture that shieds physics codes from diversity and
complexity of evolving architecturs.
\item A performant implementation of a software stack that supports this
abstraction layer.
\item Community best practices and eventual standards for this abstraction
layer.
\item Programming model/runtime must support for \gls{spmd} algorithms.
\item Programming model/runtime must support a variety of application developer data structures .
\item Programming model/runtime must support collective operations.
\item Application developer should be able to control initial problem decomposition.
\item Application developer should be able to express all form of parallelism.
\end{compactitem}


\paragraph{Co-design contributors}
In addition to the authors listed on this document,
the \gls{API} is being \gls{co-design}ed and vetted with application developers
and computer scientists whose knowledge spans the entire runtime software stack.

\paragraph{Applications affecting the design and requirements:}
\begin{compactitem}
\item Sandia \gls{ASC} \gls{ATDM} electromagnetic plasma code (POC: Matt Bettencourt)
\item Sandia \gls{ASC} \gls{ATDM} reentry code (POCs: Micah Howard, Steve Bova)
\item \gls{Trilinos} Phalanx package (POC: Roger Pawlowski)
\item Uncertainty Quantification driver (POCs: Eric Phipps, Francesco Rizzi)
\item Domain Decomposition Preconditioners (POCs: Ray Tuminaro, Clark Dohrman)
\end{compactitem}
 
\paragraph{Computer Science Research Efforts}
\begin{compactitem}
\item Kokkos (POCs: Carter Edwards, Christian Trott)
\item Data Management (POCs: Craig Ulmer, Gary Templet)
\item Low-level operating systems requirements (POCs: Stephen Olivier, Ron
    Brightwell)
\end{compactitem}



